---
title: "Assignment 1"
author: "Ayato Tanemura (atan524)"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
  
---

\newpage

```{r include=FALSE}
knitr::opts_chunk$set(comment = ">")
```
# Libraries
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse) # additional library
library(nortest)
library(car)
library(biotools)
library(PMCMRplus)
```
# Question 1
```{r include=FALSE}

df.weight <- read.csv("Data/DIET.csv") %>% 
  mutate(gender = factor(gender),
         Diet = factor(Diet),
         gender = case_when(
           gender == 0 ~ "female",
           TRUE ~ "male")
         )
weightdiff <- df.weight$pre.weight - df.weight$weight6weeks
```

## (a) Essential assumption check
### Normality Distribution

#### Normality plot:

```{r echo=FALSE, fig.align='center'}
boxplot(weightdiff ~ Diet*gender, data = df.weight)
```

* weightdiff shows the difference between before and after six weeks.

* The formula is `weightdiff = (weight before diet) - (weight 6 weeks after)`

#### Outliers check:

4 outliers are identified via this plot. 

There are 1 (Diet1, Male), 1 (Diet1, Female), 1 (Diet2, Male), and 1 (Diet3, Female) observations should be omitted. 

Therefore, 4 observations will be deleted as outliers. 

```{r include=FALSE}
# Outlier detection

outlier.weight <- boxplot(weightdiff ~ Diet*gender, data = df.weight)$out
outlier.weight

ol.weight.1.f <- df.weight[df.weight$Diet == 1 & df.weight$gender == "female" & weightdiff == 8.5,]
ol.weight.1.m <- df.weight[df.weight$Diet == 1 & df.weight$gender == "male" & weightdiff == 9.0,]
ol.weight.2.m <- df.weight[df.weight$Diet == 2 & df.weight$gender == "male" & weightdiff < 0,]
ol.weight.3.f <- df.weight[df.weight$Diet == 3 & df.weight$gender == "female" & weightdiff < 1,]

ol.weight.all <- rbind(ol.weight.1.f, ol.weight.1.m, ol.weight.2.m, ol.weight.3.f)

weight.NoOut <- df.weight[-which(df.weight$Person %in% ol.weight.all$Person),]
nrow(df.weight) - nrow(weight.NoOut)

weightdiff.NoOut = weight.NoOut$pre.weight - weight.NoOut$weight6weeks
```

#### SW test:
```{r echo=FALSE}
##SW test 
shapiro.test(weightdiff.NoOut)
```

We want to see non-significant result.

P-value for weight difference test of SW is 0.938 which is more than 0.01. H0 is not rejected and we conclude that the assumption for normality is satisfied. 

### Homogeneity of Variance

#### Levenne's test
```{r echo=FALSE}
leveneTest(weightdiff.NoOut ~ gender*Diet, data = weight.NoOut)
```

The p-value is 0.1867 which is more than 0.01 (level of significance), which means the result is non-significant.

Therefore, Homogeneity is met.


## (b) Main output & Interaction Plot

### Main output:
#### ANOVA summary:
```{r echo=FALSE}
#Assumption Check

fit.weight <- aov(weightdiff.NoOut~gender*Diet, data = weight.NoOut)
summary(fit.weight)
```

In the result of ANOVA summary, you can conclude the following, based on the p-values and a significance level of 0.01:

* The p-value for gender is 0.448, there is non-significance deference between genders in relationship with weight difference, which indicates that the gender types are not associated with volume of weight loss.

* The p-value for Diet is 0.000, there is significant differences between Diet types in relationship with weight difference, which indicates that the diet types are associated with volume of weight loss.

* The p-value for the interaction between gender*Diet is 0.000, there is significant interaction between gender and Diet, which indicates that the relationship between gender types and volume of weight loss depends on the types of diet. 

#### Tukey test:
```{r echo=FALSE}
#Tukey
TukeyHSD(fit.weight)
```

Factor gender:

Because the p-value = 0.448>0.01, it can be concluded that there is non significant difference between Male and Female. 

Factor Diet:

In the relationship with weight difference, there is not a significant difference between Diet 1 and Diet 2 (p-value = 0.764) while there are significant difference between Diet 1 and Diet 3 (p-value = 0.000); and Diet 2 and Diet 3 (p-value = 0.000).

The mean difference between Diet1 and Diet 3 is the highest.

A Tukey post hoc test revealed that the weight loss was statistically significantly increased after taking Diet 3 compared to Diet 1 and Diet 2.


### Interaction Plot
```{r echo=FALSE, fig.align='center'}
interaction.plot(weight.NoOut$Diet, weight.NoOut$gender, weightdiff.NoOut, type="b", col=c(1:3), 
                 leg.bty="o", leg.bg="beige", lwd=2, pch=c(18,24,22),	
                 xlab="Diet type", 
                 ylab="estimated marginal means", 
                 main="Interaction Plot")
```


## (c) Difference between Male Diet 1 and 2

As you can see the Tukey test result above, the p-value for male:1 - male:2 (7th row in gender:Diet table) is 0.4912294 which is larger than 0.01 significance level. The hull hypothesis is not rejected. Null hypothesis in this case is Male Diet 1 and Male Diet 2 has no difference. 

Therefore, there is no significant difference in the weight loss of males between diet 1 and diet 2. 

## (d) Highest Efficiency
As you can see from the interaction plot, Female Diet 3 has the largest impact on weight difference. 
It is regarding both gender and diet type. 

As can bee seen the difference between before and after six weeks, the larger difference of weight for Male Diet 1 and 2 while it has larger impact on weight difference for Female Diet 3. 


# Question 2
```{r}
df.stat <- read.csv("Data/STATPAK.csv") %>% 
  mutate(StatPak = factor(StatPak))
```
## (a) Outliers & Homogeneity
### Outliers

```{r fig.align='center', fig.dim = c(6, 4)}
# Outliers Detestion
olTime <- boxplot(Time ~ StatPak, data = df.stat)$out
olSatisfaction <- boxplot(Satisfaction ~ StatPak, data = df.stat)$out

olTime
olSatisfaction

out <- df.stat[df.stat$StatPak == "BMDP" & 
                 (df.stat$Time == 15 | df.stat$Time == 44 | df.stat$Time == 8),]
out
```

```{r}
# Remove Outliers
df.NoOutlier <- df.stat[-which(df.stat$No %in% out$No),] %>% 
  mutate(Satisfaction = as.numeric(Satisfaction))
```

#### SW test
```{r echo=FALSE}
# SW test for Time
shapiro.test(df.NoOutlier$Time)
```

P-value for Time of SW is 0.008 which is less than 0.01 significance level. H0 is rejected and the assumption of multivariate normality is not satisfied. 

```{r echo=FALSE}
# SW test for satisfaction
shapiro.test(df.NoOutlier$Satisfaction)
```

P-value for Satisfaction of SW is 0.03 which is more than 0.01 significance level. H0 is not rejected and the assumption of multivariate normality is satisfied. 

### Homogeneity of Variance and covariance
#### Levene's test for Time
```{r echo=FALSE}
leveneTest(Time ~ StatPak, data = df.NoOutlier)
```

P-value of Levene's for Time is 0.004 < 0.01 -> assumption is not satisfied.

Since the assumption is not met, independent group comparison would be done by Tamhane test in (c).

#### Levene's test for Satisfaction
```{r echo=FALSE}
leveneTest(Satisfaction ~ StatPak, data = df.NoOutlier)
```

P-value of Levene;s for Satisfaction is 0.6322 < 0.01 -> assumption is satisfied.

Since the assumption is met, independent group comparison would be done by Tukey test in (c).

#### Box's M test
```{r echo=FALSE}
boxM(data = df.NoOutlier[,6:7], group = df.NoOutlier$StatPak)
```

P-value of Box test is 0.007 < 0.01
-> Homogeneity of cov matrices assumption not met. We should use Pillai's Trace for interpretation of MANOVA results.

## (b) Main effects
### Pillai-Bartlett trace
```{r echo=FALSE}
# One way manova
RespoVari <- cbind(df.NoOutlier$Time,df.NoOutlier$Satisfaction)
fit.stat <- manova(RespoVari ~ StatPak, data = df.NoOutlier)
summary(fit.stat, test = 'Pillai')
```

In this case, Pillai trace is used for multivariate test. 
Based on the Pillai's Trce result, StatPak has a statistically significant effect on Time and Satisfaction as a group (F=38.6, p=0.000<0.01)

### Tests of between-subjects effects
```{r echo=FALSE}
summary.aov(fit.stat)
```
Response 1 -> Time

Response 2 -> Satisfaction

There are significant relationships between both StatPak & Time and StatPak & Satisfaction. This means different stat packages has a significant effect on the results of Time and Satisfactions as an individual. 

\newpage

## (c) Independent Group
### Tamhane test for Time
```{r echo=FALSE}
fit.T2 <- tamhaneT2Test(df.NoOutlier$Time, df.NoOutlier$StatPak)
summary(fit.T2)
```

In the relationship with Time, there is not a significant difference between SPSS and SAS(p-value=0.636>0.01) while there are significant differences between SAS and BMDP (p-vale=0.000<0.01); and SPSS and BMDP (p-value=0.000<0.01).

A Tamhane post hoc test revealed that the Time was statistically significantly decreased after taking BMDP compared to SAS and SPSS.

### Tukey test for Satisfaction
```{r echo=FALSE}
fit.TK <- aov(Satisfaction ~ StatPak, data = df.NoOutlier)
TukeyHSD(fit.TK)
```

In relationship with satisfaction, all p-value are less than .01, then there are significant difference between the independent groups.

The mean difference between SPSS and BMDP is the highest (difference value is equal to approx. 65.79).

A tukey post hoc test revealed that the satisfaction was statistically significantly increased taking SPSS compared to SAS and BMDP. Moreover, the satisfaction was statistically significantly increased SAS compared to BMDP. 
