---
title: "Assignment 1"
author: "Ayato Tanemura (atan524)"
output: 
  pdf_document:
    toc: true
    toc_depth: 4
  
---

\newpage

```{r include=FALSE}
knitr::opts_chunk$set(comment = ">")
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(nortest)
library(car)
library(RColorBrewer)
library(biotools)
library(PMCMRplus)
```
# Question 1
```{r include=FALSE}

df.weight <- read.csv("Data/DIET.csv") %>% 
  mutate(gender = factor(gender),
         Diet = factor(Diet),
         gender = case_when(
           gender == 0 ~ "female",
           TRUE ~ "male")
         )
weightdiff <- df.weight$pre.weight - df.weight$weight6weeks
```

## (a) Essential assumption check
### Normality Distribution

#### Normality plot:

```{r echo=FALSE, fig.align='center'}
boxplot(weightdiff ~ Diet*gender, data = df.weight)
```

* weightdiff shows the difference between before and after six weeks.

* The formula is `weightdiff = (weight before diet) - (weight 6 weeks after)`

#### Outliers check:

4 outliers are identified via this plot. 

There are 1 (Diet1, Male), 1 (Diet1, Female), 1 (Diet2, Male), and 1 (Diet3, Female) observations should be omitted. 

Therefore, 4 observations will be deleted as outliers. 

```{r include=FALSE}
# Outlier detection

outlier.weight <- boxplot(weightdiff ~ Diet*gender, data = df.weight)$out
outlier.weight

ol.weight.1.f <- df.weight[df.weight$Diet == 1 & df.weight$gender == "female" & weightdiff == 8.5,]
ol.weight.1.m <- df.weight[df.weight$Diet == 1 & df.weight$gender == "male" & weightdiff == 9.0,]
ol.weight.2.m <- df.weight[df.weight$Diet == 2 & df.weight$gender == "male" & weightdiff < 0,]
ol.weight.3.f <- df.weight[df.weight$Diet == 3 & df.weight$gender == "female" & weightdiff < 1,]

ol.weight.all <- rbind(ol.weight.1.f, ol.weight.1.m, ol.weight.2.m, ol.weight.3.f)

weight.NoOut <- df.weight[-which(df.weight$Person %in% ol.weight.all$Person),]
nrow(df.weight) - nrow(weight.NoOut)

weightdiff.NoOut = weight.NoOut$pre.weight - weight.NoOut$weight6weeks
```

#### SW test:
```{r echo=FALSE}
##SW test 
shapiro.test(weightdiff.NoOut)
```

We want to see non-significant result.

P-value for weight difference test of SW is 0.938 which is more than 0.05. H0 is not rejected and we conclude that the assumption for normality is satisfied. 

### Homogeneity of Variance

#### Levenne's test
```{r echo=FALSE}
leveneTest(weightdiff.NoOut ~ gender*Diet, data = weight.NoOut)
```

The p-value is more than 0.05 (level of significance), which means the result is non-significant.

Therefore, Homogeneity is met.


## (b) Interaction Plot

```{r echo=FALSE, fig.align='center'}
interaction.plot(weight.NoOut$Diet, weight.NoOut$gender, weightdiff.NoOut, type="b", col=c(1:3), 
                 leg.bty="o", leg.bg="beige", lwd=2, pch=c(18,24,22),	
                 xlab="Diet type", 
                 ylab="estimated marginal means", 
                 main="Interaction Plot")
```


## (c) Difference between Male Diet 1 and 2
### ANOVA summary:
```{r echo=FALSE}
#Assumption Check

fit.weight <- aov(weightdiff.NoOut~gender*Diet, data = weight.NoOut)
summary(fit.weight)
```

### Tukey test:
```{r echo=FALSE}
#Tamhane
summary(T2 <- tamhaneT2Test(weightdiff.NoOut, as.factor(weight.NoOut$gender) : as.factor(weight.NoOut$Diet)))
```

As you can see the Tamhane test result above, the p-value for male:1 - male:2 == 0 (7th row in gender:Diet table) is 0.9988759 which is larger than 0.05 significance level. The hull hypothesis is not rejected. Null hypothesis in this case is Male Diet 1 and Male Diet 2 has no difference. 

Therefore, there is no significant difference in the weight loss of males between diet 1 and diet 2. 

## (d) Highest Efficiency
As you can see from the interaction plot, Female Diet 3 has the largest impact on weight difference. 
It is regarding both gender and diet type. 

As can bee seen the difference between before and after six weeks, the larger difference of weight for Male Diet 1 and 2 while it has larger impact on weight difference for Female Diet 3. 


# Question 2
```{r}
df.stat <- read.csv("Data/STATPAK.csv") %>% 
  mutate(StatPak = factor(StatPak))
```
## (a) Outliers & Homogeneity
### Outliers

```{r fig.align='center', fig.dim = c(6, 4)}
# Outliers Detestion
olTime <- boxplot(Time ~ StatPak, data = df.stat)$out
olSatisfaction <- boxplot(Satisfaction ~ StatPak, data = df.stat)$out

olTime
olSatisfaction

out <- df.stat[df.stat$StatPak == "BMDP" & (df.stat$Time == 15 | df.stat$Time == 44 | df.stat$Time == 8),]
out
```

```{r}
# Remove Outliers
df.NoOutlier <- df.stat[-which(df.stat$No %in% out$No),] %>% 
  mutate(Satisfaction = as.numeric(Satisfaction))
```

#### SW test
```{r echo=FALSE}
# SW test for Time
shapiro.test(df.NoOutlier$Time)
```

P-value for Time of SW is 0.008 which is less than 0.01 significance level. H0 is rejected and the assumption of multivariate normality is not satisfied. 

```{r echo=FALSE}
# SW test for satisfaction
shapiro.test(df.NoOutlier$Satisfaction)
```

P-value for Satisfaction of SW is 0.03 which is more than 0.01 significance level. H0 is not rejected and the assumption of multivariate normality is satisfied. 

### Homogeneity of Variance and covariance
#### Levene's test for Time
```{r echo=FALSE}
leveneTest(Time ~ StatPak, data = df.NoOutlier)
```

P-value of Levene's for Time is 0.004 < 0.05 -> assumption is not satisfied.

#### Levene's test for Satisfaction
```{r echo=FALSE}
leveneTest(Satisfaction ~ StatPak, data = df.NoOutlier)
```

P-value of Levene;s for Satisfaction is 0.6322 < 0.05 -> assumption is satisfied.

#### Box's M test
```{r echo=FALSE}
boxM(data = df.NoOutlier[,6:7], group = df.NoOutlier$StatPak)
```

P-value of Box test is 0.007 < 0.05
-> Homogeneity of cov matrices assumption not met. We should use Pillai's Trace for interpretation of MANOVA results.

## (b) Main effects

```{r echo=FALSE}
# One way manova
RespoVari <- cbind(df.NoOutlier$Time,df.NoOutlier$Satisfaction)
fit.stat <- manova(RespoVari ~ StatPak, data = df.NoOutlier)
summary(fit.stat, test = 'Pillai')
```

In this case, Pillai trace is used for multivariate test. 
Based on the Pillai's Trce result, StatPak has a statistically significant effect on Time and Satisfaction as a group (F=38.6, p=0.000<0.01)

```{r echo=FALSE}
summary.aov(fit.stat)
```
Response 1 -> Time

Response 2 -> Satisfaction

There are significant relationships between both StatPak & Time and StatPak & Satisfaction. This means different stat packages has a significant effect on the results of Time and Satisfactions as an individual. 

## (c)?? Independent Group
### Tamhane test for Time
```{r echo=FALSE}
fit.T2 <- tamhaneT2Test(df.NoOutlier$Time, df.NoOutlier$StatPak)
summary(fit.T2)
```

### Tukey test for satisfaction
```{r echo=FALSE}
fit.TK <- aov(Satisfaction ~ StatPak, data = df.NoOutlier)
TukeyHSD(fit.TK)
```

